%!TEX root = ../dissertation.tex
% the acknowledgments section

This dissertation would not be possible without the support and advice of my mentors, friends, and family. First and foremost, I owe a tremendous debt of gratitude to my dissertation advisor, Chris Harvey. When I first considered choosing Chris as my advisor, I worried about joining a brand new lab with no previous students or post-docs. Over the last five years, Chris has time and time again shown me that those worries were unfounded. I have been consistently impressed with his dedication, clarity of thought, and scientific insight. It has been a privilege to learn from him, and I will miss his advice and mentorship.

\bigskip
I also want to thank all the members of the Harvey lab for our many discussions about science and life, and for their constant advice and support. Our lab environment has truly felt like one where time and advice are given without thought of return, where ideas are discussed freely, and where criticism is always constructive. So, to Selmaan Chettih, Laura Driscoll, Shin Kira, Matthias Minderer, Noah Pettit, Caroline Runyan, and Alice Wang, I thank you. I am a better scientist because of you.

\bigskip
I would also like to thank all the members of the Program in Neuroscience for creating an environment which always felt challenging, yet secure. In particular, I would like acknowledge Karen Harmin, who in addition to ensuring that the trains run on time, has also served as an invaluable friend and mentor throughout my time here. 

\bigskip
I thank the members of my Dissertation Advisory Committee --- Mark Andermann, John Assad, and Nao Uchida --- for their thoughtful advice and feedback. I also thank my Thesis Examination Committee --- Jan Drugowitsch, Mehrdad Jazayeri, and Venkatesh Murthy --- for taking the time to provide feedback on this dissertation. 

\bigskip
To all of my friends who have been with me during my time here: thank you. Thank you for complaining with me, for helping me with manuscripts and talks, and teaching me about football (baseball's still better, but football's ok, I guess). In particular, I thank Alex Trott, who has been the best lunch buddy I could have asked for. 

\bigskip
And last, but most certainly not least, I thank my girlfriend, Julia Caine, my dog, Dash, and my family for their endless love and support.  